---
title: "Capturing effects of class sizes with classroom performacne"
author: "Connor Rosenberg"
date: "1/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
#Required Packages

library("AER")
library("ggplot2")
library("car")
library("dplyr")
library("plotly")
library("MASS")

```

```{r, include=FALSE}
#Prepare Data

data("STAR")

star = data.frame( "gender" = STAR$gender, "ethnicity" = STAR$ethnicity, "star1" = STAR$star1, "math1" = STAR$math1)

star_k = data.frame( "gender" = STAR$gender, "ethnicity" = STAR$ethnicity, "star.k" = STAR$stark, "math.k" = STAR$mathk)


star = star[complete.cases(star),] #Eliminate all cases w/ NA

star_k = star_k[complete.cases(star_k),]
```

```{r}
View(star_k)
```


```{r, include=FALSE}
#Check data & objects

#str(star)

star$gender = as.factor(star$gender)
star$ethnicity = as.factor(star$ethnicity)
star$star1 = as.factor(star$star1)



```


```{r, include= FALSE}
#Set colors for pretty pictures

color2 = c("#FF6868", "#54B7FF")
color3 = c("#FF6868", "#54B7FF", "97FF68")
color4 = c("#FF6868", "#54B7FF", "97FF68", "#A168FF")
color5 = c("#FF6868", "#54B7FF", "97FF68", "#A168FF", "#FFB768")
color5 = c("#FF6868", "#54B7FF", "97FF68", "#A168FF", "#FFB768", "#FFE068")
 

#NOTE: colors will be assigned to factors in alphabetical order
```


```{r, include=FALSE}
#Task 2 Summary Statistics:
#NOTE: Please save all figures as objects to be called later in the report
#NOTE: Use plotly package to make everything consistant https://plot.ly/r

summary(star)

#Piechart of Gender
count = data.frame(table(star$gender))
pieGender = plot_ly(count, labels = ~Var1, values = ~Freq, type = 'pie') %>%
  layout(title = "Gender of Participating Studnets")
pieGender

# Piechart of star1
count2 = data.frame(table(star$star1))
pieStar1 = plot_ly(count2, labels = ~Var1, values = ~Freq, type = 'pie') %>%
  layout(title = "Class Type of 1st Grade")
pieStar1

# Bar Graph of star1

counts3 <- table(star$star1)
barplot(counts3, main="Distribution of 1st Grade Class Type", 
   xlab="Class Type", col=c("darkblue","red","green"))

#Plot of boxplot of math scores broken by class type
box1 = plot_ly(star, y = ~math1, color = ~star1, type = "box")
box1







```
From the pie chart of Gender, we can see that the numbers of males is 3414 and the number of females is 3184.




```{r, include = FALSE}
# Task 2

table(star$star1)

mean(subset(star, star1 == "small")[["math1"]])
mean(subset(star, star1 == "regular")[["math1"]])
mean(subset(star, star1 == "regular+aide")[["math1"]])

var(subset(star, star1 == "small")[["math1"]])
var(subset(star, star1 == "regular")[["math1"]])
var(subset(star, star1 == "regular+aide")[["math1"]])

```




```{r, include=FALSE}
#Task 4
```


```{r, include = FALSE}
#Fitting the model

# Test for variance homogeneity
leveneTest(math1~star1, data = star) #Equal Variacnes


#One-Way ANOVA for STAR1
reg1 = lm(math1~star1, data = star)
summary(reg1)
anova1 = anova(reg1)
anova1
boxplot(star$math1~star$star1)

plot(reg1,2)
plot(reg1, which = c(1,2))

boxcox(reg1)


#Plot of boxplot of math scores broken by class type
box1 = plot_ly(star, y = ~math1, color = ~star1, type = "box")
box1
```


```{r}
# Task 7
reg2 = lm(math.k~star.k, data = star_k)
anova2 = anova(reg2)
anova2

```

```{r, include = FALSE}
#Write the reprt below
```



# Introduction

## Background

##Questions of Interest

1. Is there a significant difference in the mean math scores in the 1st grade across different class types?
2. Are the assumptions of the ANOVA model satisfied? To explore this question, we will use model diagnostics such as residual plots and a Levene test for equality of variances.
3. Is there a significant difference in the mean math scores in kindergarten across different class types?

#Methods & Analysis

##Descriptive Analysis



